\documentclass{article}
\usepackage[utf8]{inputenc}
% %\input{GrandMacros.tex}
% %\input{Macro.tex}
\usepackage{amsthm}
\usepackage{xcolor}
\newtheorem{remark}{Remark}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{assumption}{Assumption}
\usepackage{graphicx,verbatim,array,multicol, subfigure, color, lscape, mathrsfs, dsfont,url}
%\usepackage{psfrag, amsmath, amsfonts, epsfig, fancybox, setspace,soul, cancel,biometri}
\usepackage{psfrag, amsmath, amsfonts, epsfig, fancybox, setspace, soul, amsthm, longtable, threeparttable, threeparttablex, makecell, xcolor, pdflscape, tabu, booktabs, natbib, amssymb}
\usepackage[normalem]{ulem}
\usepackage{cancel}
\usepackage{enumerate}  % enables \begin{enumerate}[(a)] labels used in the proof
\usepackage[utf8]{inputenc}
\parindent 16pt
\include{common-defs}
\newcommand{\PP}{\mathbb{P}}

\input{typography-preamble}
% \def\cor{\text{cor}}
% \def\cF{\cF}

\title{Evaluating Surrogate Transportability via Local Geometric Analysis}


\begin{document}
\maketitle

\section{Introduction}

Surrogate markers promise to accelerate clinical research by replacing expensive or long-term outcomes with earlier-measured alternatives. A central question is whether a surrogate validated in one study will perform well in future trials with different populations. This is fundamentally a question of \textbf{transportability}: will the surrogate-outcome relationship observed in the current study hold when the population, treatment effect heterogeneity, or covariate distributions change?

Traditional validation methods---proportion of treatment effect (PTE; \cite{chen2003pte}), within-study correlation, principal stratification \cite{frangakis2002principal}, and causal mediation \cite{vanderweele2015mediation}---are designed for this prospective use case. However, these methods \textbf{assume} that observed relationships will transport to future settings. PTE assumes the proportion of treatment effect remains stable; mediation assumes pathway structures persist; principal stratification assumes stratum definitions transfer. When these assumptions are violated---as may occur with covariate shift, treatment effect heterogeneity, or selection mechanisms---inference can be misleading. \cite{parast2024} identified ``limited work on transportability of surrogate knowledge from one study to another'' as a critical methodological gap.

We introduce a framework for \textbf{local geometric evaluation} of transportability. Rather than assuming the surrogate--outcome relationship transports, we characterize how a surrogate-quality functional $\phi(\cF)$ behaves across a distribution of plausible future studies $\Q \sim \cF$, restricting $\cF$ to a \emph{local geometry} $U(\PP_0, \lambda; d) = \{\Q : d(\Q, \PP_0) \leq \lambda\}$ around the observed study. The ``local'' constraint (distance $\leq \lambda$) focuses on plausible shifts; the ``geometric'' aspect (the choice of distance $d$) captures the mechanism of shift; and ``evaluation'' means we make explicit how surrogate quality varies as future studies deviate from $\PP_0$, without assuming transportability holds. Our primary instantiation takes $d$ to be total variation, $\cF$ the uniform measure on the resulting ball, and $\phi$ the correlation of treatment effects on surrogate and outcome across studies. We estimate this functional from a single study by sampling future studies via hit-and-run Markov chain Monte Carlo, reweighting the observed data to obtain each study's treatment effects, and applying a two-stage influence-function argument that yields $\sqrt{n}$-consistent, asymptotically normal inference (Section~\ref{sec:estimation}). The approach thus delivers a future-study estimand---like meta-analysis---from a single study---like mediation analysis---and applies to continuous surrogates.


\section{Setting}\label{sec:setting}
Suppose we observe $n$ iid realizations of $\bO = (\bX, A, S, Y) \sim \PP_0$ in the current study, where $A \in \{0, 1\}$ is a binary treatment, $Y \in \cY \subset \R$ is the outcome of interest, $S \in \cS \subset \R$ is a surrogate marker that may replace $Y$ in a future study, and $\bX \in \cX \subset \R^p$ are pre-treatment covariates ($\bX$ may be omitted in a randomized trial). The distribution $\PP_0$ is the law of the observables. We measure treatment effects on the surrogate and the outcome via
\begin{align*}
    &\Delta_Y(\PP_0) = \E_{\PP_0}\{Y(1) - Y(0)\}\\
    &\Delta_S(\PP_0) = \E_{\PP_0}\{S(1) - S(0)\}
\end{align*}

Under suitable identification assumptions, $\Delta_S(\PP_0)$ and $\Delta_Y(\PP_0)$ are identified and can be estimated by standard methods (e.g., difference in means or influence-function-based estimators). However, for the purposes of determining surrogate quality, we wish to know how the surrogate would perform in \emph{future} studies. Consider a future study with distribution $\Q$; we would observe (an estimate of) $\Delta_S(\Q) = \E_{\Q}\{S(1) - S(0)\}$ but not $Y$, and we wish to know what $\Delta_S(\Q)$ implies about the unobservable $\Delta_Y(\Q) = \E_{\Q}\{Y(1) - Y(0)\}$.

To formalize this, we characterize a class of future studies to which we might wish to generalize. We treat the future study distribution $\Q$ as a draw from a probability measure $\mu$ over study distributions (the ``random probability measure''), and make inference about functionals of the distribution of $\Q \sim \mu$ using data from the current study. Our primary functional is the correlation of the two treatment effects across studies,
\[
    \Theta(\mu) = \Corr_{\Q \sim \mu}\{\Delta_S(\Q), \Delta_Y(\Q)\}
    = \frac{\Cov_{\mu}\{\Delta_S(\Q), \Delta_Y(\Q)\}}{\sqrt{\Var_\mu\{\Delta_S(\Q)\}\,\Var_\mu\{\Delta_Y(\Q)\}}}.
\]
We estimate $\Theta(\mu)$ for an appropriately chosen $\mu$, thereby characterizing how the surrogate--outcome relationship may behave in future studies. Section~\ref{sec:setting} specifies $\mu$ as the uniform measure on a local geometry around $\PP_0$; we write $\phi$ for a generic functional when results apply beyond correlation.

\subsection{Local geometries for future studies}

Let $\Omega = \cX \times \{0,1\} \times \cS \times \cY$ denote the observed data space, and let $\cM_1(\Omega)$ denote the space of all probability measures on $\Omega$. The current study distribution $\PP_0$ is one element of $\cM_1(\Omega)$; future study distributions $\Q$ are also elements of this space.

We operate under the assumption that future studies are likely to bear \emph{some} resemblance to the current study -- if the current study is to be informative of the future study, they should have some similarity. We encode this notion of similarity through a \emph{distance metric} $d: \cM_1(\Omega) \times \cM_1(\Omega) \to [0,\infty)$ that quantifies how different two study distributions are. Given a distance metric $d$ and a tolerance level $\lambda \geq 0$, we define the \emph{local geometry} around $\PP_0$ as
\begin{align*}
    U(\PP_0, \lambda; d) = \{\Q \in \cM_1(\Omega) : d(\Q, \PP_0) \leq \lambda\}.
\end{align*}
This is the set of all study distributions within distance $\lambda$ of the current study. The parameter $\lambda$ controls how far future studies can deviate from $\PP_0$: when $\lambda = 0$, only $\Q = \PP_0$ is admissible; as $\lambda$ increases, more disparate future studies are considered plausible.

To characterize surrogate quality across a range of plausible future studies, we treat the future study distribution $\Q$ as random, drawn from some probability measure $\mu$ on the local geometry $U(\PP_0, \lambda; d)$. We write
\begin{align*}
    \Q \sim \mu \text{ on } U(\PP_0, \lambda; d)
\end{align*}
to denote that $\Q$ is a random draw from $\mu$ with support on $U(\PP_0, \lambda; d)$. A natural choice is the \emph{uniform measure} on $U(\PP_0, \lambda; d)$, which treats all study distributions at the same distance from $\PP_0$ as equally plausible. This uniform measure can be approximated via Markov chain Monte Carlo methods (e.g., hit-and-run sampling) when the geometry is a convex body.

The choice of $\mu$ determines which directions of deviation from $\PP_0$ are probabilistically emphasized. A uniform measure provides a conservative, non-informative assessment: surrogate quality is averaged over all directions of perturbation equally. Alternative choices (e.g., measures that concentrate near $\PP_0$) may be appropriate if prior knowledge suggests certain types of study differences are more plausible than others.

% \subsubsection{Choosing the innovation distribution $\mu$}
%
% The interpretation of surrogate quality depends critically on the choice of $\mu$, not just $\lambda$. If $\mu$ concentrates mass near $\PP_0$ (e.g., $\mu = \text{Dirichlet}(\alpha \bp_0)$ with large $\alpha$), then finding $\phi(\cF_\lambda) \geq c$ for large $\lambda$ provides only weak evidence---it merely says the surrogate performs well for future studies that closely resemble $\PP_0$. A more stringent test uses a non-informative $\mu$ that treats all innovations $\tilde{\PP} \in \cM_1(\Omega)$ equally, providing no favored direction of deviation from $\PP_0$.
%
% For computational tractability and to define a natural non-informative prior, we assume finite support:

\noindent\textbf{Assumption (Finite support).}\label{ass:finite}
The outcome space $\Omega$ has finite support. Equivalently, the observed data naturally partitions into a finite number of cells (e.g., based on observed covariate values, treatment, surrogate, and outcome levels).

This assumption is reasonable in practice: with $n$ observations, empirical distributions are inherently discrete; many applications involve categorical or discretized outcomes; and finite support makes inference tractable. Under finite support, $\cM_1(\Omega)$ is a $(k-1)$-simplex (where $k = |\Omega|$).

\subsubsection{Treatment effect heterogeneity as the fundamental object}

While the finite support assumption allows for tractable computation, it is important to recognize that the \emph{fundamental source of variation} across studies is heterogeneity in treatment effects, not arbitrary partitions of the data space. For continuous covariates $\bX$, treatment effects $\tau_S(\bX) = S(1) - S(0)$ and $\tau_Y(\bX) = Y(1) - Y(0)$ are continuous functions. Future studies differ from the current study primarily through different distributions of $\bX$, which induce different distributions of treatment effects.

The finite support assumption operationalizes this by discretizing $\bX$ into $J$ bins and treating each bin as a cell. The choice of $J$ represents an approximation: as $J \to \infty$ (with $J = o(n)$), the discretized model approaches the continuous treatment effect function. In practice, $J$ should be chosen large enough to capture meaningful heterogeneity but small enough to avoid over-discretization (which increases estimation noise).

This perspective clarifies that the fundamental object is the continuous treatment-effect heterogeneity, not a set of discrete latent classes. Discretization schemes that align with this heterogeneity (e.g., partitioning on estimated $\tau(\bX)$) approximate it more faithfully than arbitrary covariate-based partitions.

% and we can define a canonical non-informative prior.
%
% Under finite support, the natural choice for a non-informative $\mu$ is the uniform Dirichlet:
% \begin{align}
%     \mu = \text{Dirichlet}(1, \ldots, 1).
% \end{align}
% This distribution is uniform on the simplex $\cM_1(\Omega)$, treating all probability vectors equally. It is the maximum entropy prior given no constraints, providing the most conservative (stringent) assessment of surrogate quality. When $\phi(\cF_\lambda) \geq c$ under this choice of $\mu$, the surrogate performs well across a broad, evenly-distributed set of future studies within distance $\lambda$ of $\PP_0$.
%
% \subsubsection{How similar must a future study be for the surrogate to remain good?}
%
% A central question is: \emph{how different can a future study be from the current study while the surrogate remains good?} We formalize this by asking for what values of $\lambda$ the functional $\phi(\cF_\lambda)$ stays above a desired threshold. For example, if we define ``good surrogate'' as correlation of treatment effects above 0.8, we want to find the largest $\lambda$ for which $\phi(\cF_\lambda) \geq 0.8$. This $\lambda$ tells us how far a future study can deviate from the current study (in total variation distance) while maintaining adequate surrogate quality.
%
% To make this operational, we need to understand how functionals $\phi(\cF_\lambda)$ behave as $\lambda$ varies. The key insight is that the treatment effects $\Delta_S(\Q)$ and $\Delta_Y(\Q)$ depend continuously on $\Q$ (in the sense that small changes in $\Q$ produce small changes in the treatment effects), and thus functionals of the treatment effects also vary smoothly with $\lambda$. The following assumptions and result formalize this.
%
% \noindent\textbf{Assumption 3 (Bounded outcomes).}
% $S \in [s_{\min}, s_{\max}]$ and $Y \in [y_{\min}, y_{\max}]$ are bounded.
%
% Under this assumption, treatment effects are Lipschitz continuous in $\Q$. For any $\Q, \Q'$,
% \begin{align}\label{eq:lipschitz}
%     |\Delta_S(\Q) - \Delta_S(\Q')| &\leq 2(s_{\max} - s_{\min}) \, d_{\mathrm{TV}}(\Q, \Q')\\
%     |\Delta_Y(\Q) - \Delta_Y(\Q')| &\leq 2(y_{\max} - y_{\min}) \, d_{\mathrm{TV}}(\Q, \Q'). \nonumber
% \end{align}
% This follows because expectations of bounded functions are Lipschitz in total variation: $|\E_{\Q} f - \E_{\Q'} f| \leq (\sup f - \inf f) \, d_{\mathrm{TV}}(\Q, \Q')$. Under SUTVA and consistency, $S(1) - S(0)$ ranges over $[-(s_{\max} - s_{\min}), (s_{\max} - s_{\min})]$, yielding the constant $2(s_{\max} - s_{\min})$, and similarly for $Y$.
%
% \begin{theorem}[Surrogate quality under perturbations]\label{thm:epsilon}
% Suppose ``good surrogate'' is defined by $\phi(\cF_\lambda) \geq c$ for some threshold $c$ (e.g., $\phi$ is the correlation of treatment effects and $c = 0.8$). Under Assumption 3, the functional $\phi(\cF_\lambda)$ varies continuously with $\lambda$. If $\phi(\cF_\lambda) \geq c$ for some $\lambda$, then any future study $\Q$ with $d_{\mathrm{TV}}(\Q, \PP_0) \leq \lambda$ will have $|\phi(\Q) - \phi(\cF_\lambda)| \leq L \lambda$ for some constant $L$ depending on $\phi$ and regularity conditions (e.g., non-degenerate variances for correlation).
% \end{theorem}
% \begin{proof}
% By \eqref{eq:lipschitz}, the treatment effects $(\Delta_S(\Q), \Delta_Y(\Q))$ are Lipschitz continuous in $\Q$ with respect to $d_{\mathrm{TV}}$. The functionals $\phi$ of interest (correlation, probability of concordance, conditional expectation) are continuous---and under bounded $S$, $Y$ and non-degenerate variances, Lipschitz---as functions of the bivariate distribution of treatment effects. Since $\Q \in \cF_\lambda(\PP_0)$ implies $d_{\mathrm{TV}}(\Q, \PP_0) \leq \lambda$ by \eqref{eq:distance-bound}, the distribution of $(\Delta_S(\Q), \Delta_Y(\Q))$ varies continuously with $\lambda$, and composing the Lipschitz maps yields continuity of $\phi(\cF_\lambda)$ in $\lambda$.
% \end{proof}
%
% \begin{remark}[Operationalizing the threshold]\label{remark:operational}
% In practice, we estimate $\phi(\cF_\lambda)$ for a range of $\lambda$ values (e.g., $\lambda \in \{0.05, 0.1, 0.15, \ldots, 0.5\}$) by generating future studies $\Q = (1-\lambda)\hat{\PP}_n + \lambda\tilde{\PP}$ with $\tilde{\PP} \sim \mu$ (e.g., $\mu = \text{Dirichlet}(1, \ldots, 1)$), computing $(\Delta_S(\Q), \Delta_Y(\Q))$ for each, and calculating the empirical correlation (or other functional). The largest $\lambda$ for which $\hat{\phi}(\cF_\lambda) \geq c$ provides an assessment: \emph{when future studies are drawn from $\cF_\lambda$ (within distance $\lambda$ and distributed according to $\mu$), the surrogate performs well on average.} This operationalizes ``how different future studies can be while maintaining adequate surrogate quality.''
%
% Importantly, $\phi(\cF_\lambda) \geq c$ is a statement about the \emph{average} surrogate performance over the distribution $\mu$, not a guarantee for every individual study within distance $\lambda$. With uniform $\mu = \text{Dirichlet}(1, \ldots, 1)$, this average represents a stringent, conservative test.
% \end{remark}
%
% \begin{remark}[Sensitivity to $\mu$]\label{remark:mu-sensitivity}
% We recommend reporting $\hat{\phi}(\cF_\lambda)$ under multiple choices of $\mu$ to assess sensitivity. For example:
% \begin{itemize}
%     \item $\mu_{\text{uniform}} = \text{Dirichlet}(1, \ldots, 1)$ (non-informative, most stringent)
%     \item $\mu_{\text{moderate}} = \text{Dirichlet}(\alpha \bp_0)$ with moderate $\alpha$ (e.g., $\alpha = 5$)
%     \item $\mu_{\text{concentrated}} = \text{Dirichlet}(\alpha \bp_0)$ with large $\alpha$ (e.g., $\alpha = 50$)
% \end{itemize}
% If $\hat{\phi}(\cF_\lambda) \geq c$ for large $\lambda$ even under $\mu_{\text{uniform}}$, this provides strong evidence of robust surrogate quality. If this holds only under $\mu_{\text{concentrated}}$, the evidence is weaker (surrogate is good only for futures resembling $\PP_0$).
% \end{remark}

\subsection{Estimation}\label{sec:estimation}

\subsubsection{Setup for fixed $\lambda$}

For a fixed value $\lambda \geq 0$ and distance metric $d$, we define the local geometry
$U(\PP_0, \lambda; d)$ as in Section~\ref{sec:setting}. The estimand is the functional
$\phi$ evaluated over a probability measure $\mu$ on this local geometry. We treat $\lambda$
as a design parameter—analogous to choosing a significance level—that
controls how far future studies may deviate from $\PP_0$.

In this section, we develop asymptotic theory for estimating $\phi(\mu)$
for a single fixed $\lambda$ and $d$. In practice one reports
$\hat{\phi}(\mu)$ across a grid of $\lambda$ values to assess how surrogate
quality varies with the assumed magnitude of distributional shift
(Section~\ref{sec:sim-future}).

\subsubsection{Regularity conditions}
We assume the following.

\noindent\textbf{Assumption 1 (Bounded outcomes).}\label{ass:bounded}
$S \in [s_{\min}, s_{\max}]$ and $Y \in [y_{\min}, y_{\max}]$ almost surely
under $\PP_0$ and all $\Q \in U(\PP_0, \lambda; d)$, with
$0 < s_{\max} - s_{\min} < \infty$ and $0 < y_{\max} - y_{\min} < \infty$.

\noindent\textbf{Assumption 2 (Identification and estimability).}\label{ass:identification}
The treatment effects $\Delta_S(\Q)$ and $\Delta_Y(\Q)$ are identified from
observables under standard causal identification conditions (e.g., randomization,
unconfoundedness, or instrumental variables), and admit $\sqrt{n}$-consistent
estimators with influence functions satisfying $\E[\psi^2] < \infty$.

\noindent\textbf{Assumption 3 (Functional smoothness).}\label{ass:smoothness}
For the correlation functional, write $\Theta = \varphi(m)$ as a function of the
vector of moments
$m = (\E_\mu \Delta_S, \E_\mu \Delta_Y, \E_\mu \Delta_S^2, \E_\mu \Delta_Y^2, \E_\mu \Delta_S\Delta_Y)$
of the treatment-effect pair $(\Delta_S(\Q), \Delta_Y(\Q))$ over $\Q \sim \mu$ on
$U(\PP_0, \lambda; d)$. The map $\varphi$ is continuously differentiable with
bounded gradient $\|\nabla\varphi(m)\| < \infty$ on the region where the
across-study variances are bounded away from zero (Assumption~\ref{ass:nondegeneracy}).

\noindent\textbf{Assumption 4 (Non-degeneracy).}\label{ass:nondegeneracy}
For the correlation functional, there exists $\epsilon > 0$ such that
\begin{align*}
    \Var_\mu[\Delta_S(\Q)] &\geq \epsilon\\
    \Var_\mu[\Delta_Y(\Q)] &\geq \epsilon
\end{align*}
where the variance is taken over $\Q \sim \mu$ on $U(\PP_0, \lambda; d)$
(i.e., variation across different future studies, not within a single study).

\begin{remark}[On finite support]
Assumption~\ref{ass:finite} (finite support, stated in Section~\ref{sec:setting}) is restrictive
and primarily computational. It renders the local geometry a convex body in the
probability simplex, so the uniform measure $\mu$ on $U(\PP_0, \lambda; d)$ can be
sampled by hit-and-run MCMC. In practice, continuous outcomes can be
discretized into a fine grid (trading bias for computational feasibility), or
replaced by kernel-based methods (left for future work). The asymptotic theory
extends to growing support $k=k_n$ under additional rate conditions, but we do not
pursue that here.
\end{remark}

Under Assumption~\ref{ass:finite}, the empirical distribution $\hat{\PP}_n$ is the vector of empirical proportions $\hat{\bp}_n$; $\sqrt{n}(\hat{\bp}_n - \bp_0)$ is asymptotically normal (multinomial).

\subsubsection{Computational approach: Uniform sampling via hit-and-run MCMC}

For many local geometries (TV ball, chi-squared ball, $L_2$ ball), $U(\PP_0, \lambda; d)$ forms a convex body in the probability simplex. To approximate the uniform measure $\mu$ on this geometry, we use \emph{hit-and-run sampling}, a Markov chain Monte Carlo method for uniformly sampling from convex bodies.

\paragraph{Hit-and-run algorithm.}
Given current state $\Q_t \in U(\PP_0, \lambda; d)$:
\begin{enumerate}
    \item Draw a random direction $\bv \in \R^k$ (e.g., from standard Gaussian)
    \item Compute the line segment $L_t = \{\Q_t + \alpha \bv : \alpha \in \R, \, \Q_t + \alpha \bv \in U(\PP_0, \lambda; d)\}$
    \item Sample $\Q_{t+1}$ uniformly along $L_t$
\end{enumerate}
After a burn-in period, the chain converges to the uniform distribution on $U(\PP_0, \lambda; d)$.

\paragraph{The plug-in estimator.}
For fixed $\lambda$ and distance metric $d$, define the plug-in estimator
\[
    \hat{\phi}_n(\lambda) = \frac{1}{M}\sum_{m=1}^M \phi(\Q_m)
\]
where $\Q_1, \ldots, \Q_M$ are drawn (after burn-in) from hit-and-run MCMC on $U(\hat{\PP}_n, \lambda; d)$. Operationally, we:
\begin{enumerate}
    \item Run hit-and-run MCMC to generate $M$ samples $\Q_m$ from $U(\hat{\PP}_n, \lambda; d)$
    \item For each $\Q_m$, compute treatment effects $(\hat{\Delta}_S(\Q_m), \hat{\Delta}_Y(\Q_m))$
          via deterministic reweighting of observed data
    \item Compute $\hat{\phi}_n(\lambda)$ as the sample correlation (or other functional)
          of the $M$ pairs $\{(\hat{\Delta}_S(\Q_m), \hat{\Delta}_Y(\Q_m))\}_{m=1}^M$
\end{enumerate}

\paragraph{Treatment effect estimation via deterministic reweighting.}
For each $\Q_m$ drawn from the geometry, we do \emph{not} resample data. Instead, we compute observation weights $w_i^{(m)} = q_m(O_i)$ where $q_m$ is the probability mass function of $\Q_m$. The treatment effects are then estimated by applying standard causal inference methods to the weighted data.

\textit{Randomized trials.} Under randomization, treatment effects are estimated as weighted differences in means:
\begin{align*}
    \hat{\Delta}_S(\Q_m) &= \frac{\sum_i w_i^{(m)} A_i S_i}{\sum_i w_i^{(m)} A_i} - \frac{\sum_i w_i^{(m)} (1-A_i) S_i}{\sum_i w_i^{(m)} (1-A_i)}\\
    \hat{\Delta}_Y(\Q_m) &= \frac{\sum_i w_i^{(m)} A_i Y_i}{\sum_i w_i^{(m)} A_i} - \frac{\sum_i w_i^{(m)} (1-A_i) Y_i}{\sum_i w_i^{(m)} (1-A_i)}.
\end{align*}

\textit{Observational studies.} Under unconfoundedness (selection on observables), treatment effects are estimated using augmented inverse probability weighting (AIPW) with cross-fitting to avoid overfitting bias. Let $\hat{e}(\bX_i)$ denote the estimated propensity score and $\hat{\mu}_a(\bX_i)$ the estimated outcome regression for treatment $a \in \{0,1\}$, both fit using cross-validation folds. The doubly-robust AIPW estimator is:
\begin{align*}
    \hat{\Delta}_S(\Q_m) &= \sum_i w_i^{(m)} \left[\frac{A_i(S_i - \hat{\mu}_1^S(\bX_i))}{\hat{e}(\bX_i)} - \frac{(1-A_i)(S_i - \hat{\mu}_0^S(\bX_i))}{1-\hat{e}(\bX_i)} + \hat{\mu}_1^S(\bX_i) - \hat{\mu}_0^S(\bX_i)\right]\\
    \hat{\Delta}_Y(\Q_m) &= \sum_i w_i^{(m)} \left[\frac{A_i(Y_i - \hat{\mu}_1^Y(\bX_i))}{\hat{e}(\bX_i)} - \frac{(1-A_i)(Y_i - \hat{\mu}_0^Y(\bX_i))}{1-\hat{e}(\bX_i)} + \hat{\mu}_1^Y(\bX_i) - \hat{\mu}_0^Y(\bX_i)\right].
\end{align*}
Cross-fitting ensures that nuisance parameters ($\hat{e}, \hat{\mu}_a$) are estimated on independent data, enabling asymptotic normality under weak conditions (Chernozhukov et al., 2018).

In both cases, the deterministic reweighting approach eliminates sampling variability within each $\Q_m$, leaving only two sources of error: estimation of $\PP_0$ and Monte Carlo approximation of the geometry.

\paragraph{Variance decomposition.}
The variance of $\hat{\phi}_n(\lambda)$ arises from two sources:
\begin{enumerate}
    \item \textbf{Estimation error:} The empirical distribution $\hat{\PP}_n$ differs from $\PP_0$ by $O_p(n^{-1/2})$, and in observational studies, nuisance parameters (propensity scores, outcome regressions) are estimated with error. Under cross-fitting, this contributes $O_p(n^{-1/2})$ to the overall rate.
    \item \textbf{Monte Carlo error from MCMC:} We use $M$ draws from hit-and-run
          to approximate the uniform expectation $\E_{\text{Uniform}}[\phi(\Q)]$ over
          $U(\PP_0, \lambda; d)$, contributing error $O_p(M^{-1/2})$.
\end{enumerate}

Under Assumptions~\ref{ass:bounded}--\ref{ass:smoothness}, the dominant term is
(1) when $M = o(n)$ and MCMC has converged. Thus, $M = O(n^{1/2})$ suffices to make Monte Carlo error negligible compared to sampling error.

\paragraph{Asymptotic normality.}
The population estimand $\Theta(\PP_0, \lambda)$ is the functional evaluated over the
uniform measure $\mu$ on the local geometry; for the correlation functional,
\[
    \Theta(\PP_0, \lambda) = \Corr_{\Q \sim \mu}\{\Delta_S(\Q), \Delta_Y(\Q)\}
    = \frac{\Cov_\mu\{\Delta_S(\Q), \Delta_Y(\Q)\}}{\sqrt{\Var_\mu\{\Delta_S(\Q)\}\,\Var_\mu\{\Delta_Y(\Q)\}}},
\]
a single functional of the distribution over $\Q \sim \mu$ (not an average of a
per-study quantity), consistently estimated by the sample correlation of the $M$
treatment-effect pairs $\{(\hat\Delta_S(\Q_m), \hat\Delta_Y(\Q_m))\}_{m=1}^M$.

\begin{proposition}\label{prop:asymptotic}
Suppose Assumptions~\ref{ass:bounded}--\ref{ass:nondegeneracy} hold. For fixed $\lambda$ and $d$,
conditional on the $M$ sampled studies and after MCMC burn-in,
\[
    \sqrt{n}\{\hat{\Theta}_n(\lambda) - \Theta(\PP_0, \lambda)\}
    \xrightarrow{d} N(0, \sigma^2(\lambda))
\]
where $\sigma^2(\lambda) = \E[\psi_\Theta(O; \PP_0)^2]$ for an influence function $\psi_\Theta$ that accounts for treatment effect estimation error. The Monte Carlo error from finite $M$ is $O_p(M^{-1/2})$ and is negligible relative to the $\sqrt{n}$ term when $M$ grows faster than $n$; otherwise inference is conditional on the sampled studies. A complete statement, including the two regimes, appears in the supplementary material.
\end{proposition}

\begin{proof}[Proof sketch]
The proof proceeds in two stages, corresponding to the two-stage structure of the estimator.

\textit{Stage 1: Treatment effects.} For each $\Q \in U(\PP_0, \lambda; d)$, the treatment effects $\Delta_S(\Q)$ and $\Delta_Y(\Q)$ are estimated via deterministic reweighting. In RCTs with treatment probability $\pi = \PP_0(A=1)$, the weighted difference in means has influence function
\[
    \psi_{\Delta_S}(O; \Q) = w(\Q, O) \left[\frac{A\,(S - m_{S,1})}{\pi} - \frac{(1-A)\,(S - m_{S,0})}{1-\pi}\right],
\]
where $w(\Q, O) = q(O)/p_0(O)$ is the importance weight and $m_{S,a}$ is the (weighted) mean of $S$ in arm $a$. In observational studies with cross-fitted AIPW, the influence function takes the doubly-robust form that removes nuisance parameter estimation bias \cite{chernozhukov2018}.

\textit{Stage 2: Functional aggregation.} For the correlation functional, $\Theta$ is a smooth function of five moments of the distribution of $(\Delta_S(\Q), \Delta_Y(\Q))$ over $\Q \sim \mu$ ($\E_\mu \Delta_S$, $\E_\mu \Delta_Y$, $\E_\mu \Delta_S^2$, $\E_\mu \Delta_Y^2$, $\E_\mu \Delta_S\Delta_Y$). By Hadamard differentiability (Assumption~\ref{ass:smoothness}, with the non-degeneracy of Assumption~\ref{ass:nondegeneracy}), the functional delta method composes the Stage-1 influence functions with the moment gradient to yield the overall influence function $\psi_\Theta$, giving asymptotic normality.

A complete derivation is provided in the supplementary material, for both randomized and observational studies. For observational studies with cross-fitted AIPW, the key property is Neyman orthogonality under reweighting, which combined with $o_P(n^{-1/4})$ nuisance convergence rates yields an $o_P(n^{-1/2})$ remainder uniformly over the sampled studies. The Monte Carlo term is handled by a Markov chain central limit theorem for the geometrically ergodic hit-and-run sampler.
\end{proof}

\paragraph{Bootstrap inference.}
Direct computation of the asymptotic variance requires explicit influence function formulas and may be
unstable in small samples (especially with cross-fitting in observational studies). We recommend the bootstrap for practical inference:
\begin{enumerate}
    \item For $b = 1, \ldots, B$ bootstrap replicates:
          \begin{enumerate}
              \item Resample the data with replacement to get bootstrap sample $\{O_i^{(b)}\}_{i=1}^n$
              \item Fit nuisance parameters (propensity scores, outcome regressions) on bootstrap sample if using AIPW
              \item Run hit-and-run MCMC to generate $M$ draws $\{\Q_m^{(b)}\}$ from $U(\hat{\PP}_n^{(b)}, \lambda; d)$
              \item Compute $\hat{\phi}_n^{(b)}(\lambda)$ via deterministic reweighting
          \end{enumerate}
    \item Use the bootstrap distribution $\{\hat{\phi}_n^{(b)}(\lambda)\}_{b=1}^B$
          for inference (percentile intervals or standard errors)
\end{enumerate}

\begin{remark}[Bootstrap validity]
The bootstrap is valid under the conditions of Proposition~\ref{prop:asymptotic}. In RCTs, the bootstrap correctly mimics the sampling distribution of weighted means. In observational studies with cross-fitting, refitting nuisance parameters on each bootstrap sample preserves the doubly-robust structure. Standard bootstrap theory for smooth functionals applies.
\end{remark}

\section{Example: TV-Ball Geometry}\label{sec:tv-implementation}

We now instantiate the local geometric evaluation framework using total variation distance. The TV-ball
\begin{equation}
    U(\PP_0, \lambda; d_{\mathrm{TV}}) = \{\Q : d_{\mathrm{TV}}(\Q, \PP_0) \leq \lambda\}
\end{equation}
captures arbitrary distributional shifts---any reweighting of the observed data that changes probabilities by at most $\lambda$. This is the most general local geometry, appropriate when the mechanism of distributional shift is unknown or when maximal robustness is desired.

As described in Section~\ref{sec:estimation}, we characterize surrogate quality over this geometry by:
\begin{enumerate}
    \item Using hit-and-run MCMC to sample uniformly from $U(\PP_0, \lambda; d_{\mathrm{TV}})$
    \item Computing treatment effects $(\Delta_S(\Q), \Delta_Y(\Q))$ for each sampled $\Q$ via deterministic reweighting
    \item Estimating the functional $\Theta(\PP_0) = \E_{\text{Uniform}}[\phi(\Q)]$ as the sample mean over MCMC draws
\end{enumerate}

This section provides implementation details and computational considerations specific to the TV-ball geometry.

\subsection{Hit-and-run sampling for TV balls}

For the TV-ball $U(\PP_0, \lambda; d_{\mathrm{TV}})$, the hit-and-run MCMC algorithm proceeds as follows. Given current state $\Q_t \in U(\PP_0, \lambda; d_{\mathrm{TV}})$:

\begin{enumerate}
    \item \textbf{Draw random direction:} Sample $\bv \sim N(0, I_k)$ on the probability simplex
    \item \textbf{Compute line segment:} Find the interval $[\alpha_{\min}, \alpha_{\max}]$ such that $\Q_t + \alpha \bv \in U(\PP_0, \lambda; d_{\mathrm{TV}})$ for all $\alpha \in [\alpha_{\min}, \alpha_{\max}]$. This requires solving:
    \begin{align*}
        d_{\mathrm{TV}}(\Q_t + \alpha \bv, \PP_0) &\leq \lambda\\
        \sum_{i=1}^k |q_{t,i} + \alpha v_i - p_{0,i}| / 2 &\leq \lambda
    \end{align*}
    \item \textbf{Sample along line:} Draw $\alpha \sim \text{Uniform}([\alpha_{\min}, \alpha_{\max}])$ and set $\Q_{t+1} = \Q_t + \alpha \bv$
    \item \textbf{Project to simplex:} If $\Q_{t+1}$ violates simplex constraints (negative entries or doesn't sum to 1), project back
\end{enumerate}

After a burn-in period (typically $B = 100$ iterations), the chain converges to the uniform distribution on $U(\PP_0, \lambda; d_{\mathrm{TV}})$. For implementation details and convergence diagnostics, see the R package documentation.

\paragraph{Computational considerations.}
The key computational bottleneck is solving for $[\alpha_{\min}, \alpha_{\max}]$ at each iteration, which requires checking TV-distance constraints. For finite support with $k$ cells, this is $O(k)$ per iteration. With $M = 1000$ MCMC draws and $B = 100$ burn-in, total cost is approximately $(M + B) \times k$ distance evaluations plus treatment effect computations on reweighted data.

For moderate sample sizes ($n = 500$, $k = 50$ cells), hit-and-run with $M = 1000$ takes 1-2 seconds on standard hardware. The method scales linearly in $k$, making it practical for problems with tens to hundreds of cells (corresponding to discretized covariate spaces).

Unlike Dirichlet-based sampling of the simplex---whose concentration around $\PP_0$ depends on a scale parameter and which does not yield the uniform measure on a distance ball---hit-and-run targets the uniform measure on $U(\PP_0, \lambda; d_{\mathrm{TV}})$ directly, giving each direction of deviation equal weight.

\section{Numerical studies}\label{sec:simulation}

We validate the asymptotic theory of Section~\ref{sec:estimation} and illustrate
how the across-study correlation behaves in settings where the proportion of
treatment effect (PTE) is misleading. All studies use the TV-ball geometry with
the correlation functional and the importance-weighted estimator for randomized
trials.

\subsection{Design}\label{sec:sim-design}

We use four data-generating processes with a common structure and a single
five-level categorical covariate $X \in \{-2,-1,0,1,2\}$ drawn with probabilities
$(0.05, 0.25, 0.40, 0.25, 0.05)$. Conditional on $X$ and treatment $A \sim
\mathrm{Bernoulli}(0.5)$,
\begin{align*}
    S &= (\gamma_A + \gamma_{AX} X)\,A + \epsilon_S, & \epsilon_S &\sim N(0, 0.5^2),\\
    Y &= (\beta_A + \beta_{AX} X)\,A + \beta_S S + \beta_{SX} (S X) + \epsilon_Y,
        & \epsilon_Y &\sim N(0, 0.5^2).
\end{align*}
The four DGPs vary the coefficients to produce different combinations of PTE and
across-study correlation (Table~\ref{tab:dgps}). For each replication we use a
sample size of $n = 10{,}000$ (deliberately large, to isolate bias from variance),
draw $M \approx 2100$ future studies from the TV ball of radius $\lambda = 0.3$
via adaptive hit-and-run MCMC, estimate the correlation $\hat{\Theta}_n(0.3)$ of
treatment effects across studies, and form influence-function-based confidence
intervals. We run 1000 replications per DGP. The true across-study correlation
$\Theta$ for each DGP is computed analytically from the cell-level treatment
effects and reported in Table~\ref{tab:dgps}.

\begin{table}[h]
\centering
\caption{Data-generating processes. Coefficients as coded in the
\texttt{surrogateTransportability} package (\texttt{canonical\_dgp\_params}); true
across-study correlation $\Theta$ and PTE at $\PP_0$ computed analytically.}
\label{tab:dgps}
\begin{tabular}{lcccccccc}
\toprule
DGP & $\gamma_A$ & $\gamma_{AX}$ & $\beta_A$ & $\beta_{AX}$ & $\beta_S$ & $\beta_{SX}$ & $\Theta$ & PTE \\
\midrule
1 & 1.0 & 0.5 & 0.25 & $-0.3$ & 0.9 & $-0.10$ & $0.69$   & 0.82 \\
2 & 1.0 & 0.5 & 0.60 & $-0.3$ & 0.6 & $-0.15$ & $-0.88$  & 0.53 \\
3 & 1.0 & 0.5 & 0.70 & $0.4$  & 0.3 & $0.00$  & $1.00$   & 0.30 \\
4 & 0.0 & 0.5 & 0.00 & $0.5$  & 0.6 & $0.00$  & $1.00$   & undef. \\
\bottomrule
\end{tabular}
\end{table}

DGP~1 has high mediation and moderate positive correlation (a baseline where PTE
and correlation broadly agree). DGP~2 differs from DGP~1 mainly through a stronger
negative surrogate--covariate interaction ($\beta_{SX} = -0.15$) and weaker direct
mediation ($\beta_S = 0.6$), which together drive a strong \emph{negative}
across-study correlation despite moderate PTE. DGP~3 has weak mediation ($\beta_S = 0.3$, PTE
$=0.30$) but effect modification that scales both effects together, giving
near-perfect correlation---low PTE does not imply a poor surrogate. DGP~4 has
antisymmetric treatment effects and a symmetric $\PP_0$, so $\Delta_Y(\PP_0)
\approx 0$ and PTE is undefined, while the across-study correlation remains
well-defined and near one.

\subsection{Results}\label{sec:sim-results}

\begin{table}[h]
\centering
\caption{Finite-sample performance over 1000 replications ($n = 10{,}000$,
$\lambda = 0.3$). \textbf{[PENDING: values to be populated from the O2 cluster
re-validation (\texttt{simulations/canonical-validation/}) run with the corrected
estimator; the pre-correction run is not reported here.]}}
\label{tab:performance}
\begin{tabular}{lccccc}
\toprule
DGP & $\Theta$ (true) & Bias & Emp.\ SE & Est.\ SE & Coverage \\
\midrule
1 & $0.69$  & --- & --- & --- & --- \\
2 & $-0.88$ & --- & --- & --- & --- \\
3 & $1.00$  & --- & --- & --- & --- \\
4 & $1.00$  & --- & --- & --- & --- \\
\bottomrule
\end{tabular}
\end{table}

Figures and Table~\ref{tab:performance} are generated from the cluster
re-validation output. We report bias, empirical and estimated standard errors,
and coverage of nominal 95\% confidence intervals for each DGP; the influence-
function standard error is expected to track the empirical standard deviation and
coverage to be at or near nominal, validating the two-stage theory of
Section~\ref{sec:estimation}.

\subsection{When PTE misleads}\label{sec:pte-misleads}

The four DGPs show three regimes in which PTE and the across-study correlation
diverge. In DGP~2, PTE is moderate ($0.53$) and would not obviously disqualify the
surrogate, yet the across-study correlation is strongly \emph{negative} ($-0.88$):
treatment effects move in opposite directions across populations, so the surrogate
does not transport. In DGP~3, low PTE ($0.30$) coexists with near-perfect
transportability---low mediation does not imply a poor surrogate. In DGP~4, PTE is
undefined ($\Delta_Y(\PP_0) \approx 0$) yet the correlation is well-defined and
near one. In each case mediation-based reasoning is either misleading or
inapplicable, while the local geometric evaluation returns a well-defined
transportability assessment. More generally, opposite-signed effect modification
on surrogate and outcome can make PTE at $\PP_0$ look favorable while treatment
effects are nearly uncorrelated across studies; such a configuration is a natural
target for the finite-sample studies in Section~\ref{sec:sim-future}.

\subsection{Further numerical studies (in progress)}\label{sec:sim-future}

The studies above fix $n = 10{,}000$ and $\lambda = 0.3$ and use the total-variation
geometry, to validate the asymptotic theory in the regime where it is cleanest.
Three extensions are planned and reported in the supplementary material as they
complete. First, \textbf{finite-sample behavior at practical sample sizes}
($n \in \{250, 500, 1000, 2500, 5000\}$): at smaller $n$ the correlation estimate
is attenuated and the normal approximation degrades, and we characterize the
resulting coverage and study weight-stabilization remedies for the observational
(AIPW) path. Second, \textbf{alternative geometries}: repeating the validation
under $\chi^2$ and $L_2$ balls to compare which distributional shifts each
geometry emphasizes. Third, \textbf{sensitivity to $\lambda$}: reporting
$\hat{\Theta}_n(\lambda)$ across a grid of $\lambda$ values, where a flat profile
indicates robust transportability and a declining profile indicates fragility.

\section{Discussion}\label{sec:discussion}

We introduced a framework for evaluating surrogate transportability from a single
study by modeling hypothetical future studies as random probability measures
within a local geometry around the observed study, and estimating a functional---
principally the correlation of treatment effects on surrogate and outcome---across
those studies. The estimator is $\sqrt{n}$-consistent and asymptotically normal,
with influence-function-based inference, and extends to observational studies via
cross-fitted augmented inverse probability weighting. The framework thus provides
a future-study estimand, like meta-analysis, from a single study, like mediation
analysis, and applies to continuous surrogates.

\paragraph{Transportability is distinct from within-study mediation.}
The correlation $\Theta(\PP_0, \lambda)$ measures whether treatment effects
co-vary across plausible future studies; the proportion of treatment effect
measures how much of the effect is mediated through the surrogate in the current
study. These are different questions, and the numerical studies show they can
disagree sharply: a surrogate may exhibit weak mediation yet perfect
transportability (DGP~3), or PTE may be undefined while the correlation is
well-defined and near one (DGP~4). A surrogate that looks good by PTE can still
fail to transport when effect modification acts in opposite directions on
surrogate and outcome.

\paragraph{Interpreting $\lambda$.}
The radius $\lambda$ encodes how far future studies may plausibly deviate from
$\PP_0$ in total variation. Rather than commit to a single value, we recommend
reporting $\hat{\Theta}_n(\lambda)$ across a grid: a flat profile indicates
transportability that is robust to substantial distributional shift, while a
declining profile flags fragility. Data-driven or decision-theoretic choices of
$\lambda$ are a natural direction for future work.

\paragraph{Limitations.}
The absolute-continuity restriction $\Q \ll \PP_0$ avoids extrapolation but
confines attention to reweightings of the observed covariate support; relaxing it
to allow genuinely new covariate values requires stronger modeling assumptions.
The finite-support assumption is computationally convenient and requires
discretizing continuous covariates; the across-study variances must be
non-degenerate for the correlation functional to be well-defined. Finite-sample
behavior at small $n$, alternative geometries, and $\lambda$-sensitivity are under
active study (Section~\ref{sec:sim-future}).

\paragraph{Extensions.}
The framework accommodates other smooth functionals beyond correlation---for
example $R^2$ (variance explained) and mean squared prediction error. It applies
directly to prediction-powered inference, where one asks whether a machine-learning
prediction is a reliable surrogate for an expensive gold-standard label across
deployment contexts. Software implementing the framework is available in the R
package \texttt{surrogateTransportability}.

  \input{proof_asymptotic_normality}

\bibliography{refs}
\bibliographystyle{plain}
\end{document}
